---
title: "O2 LRC Fitting"
output: html_notebook
author: Douglas A. Campbell
---

This R Notebook imports oxygen evolution vs. PAR data.
It fits and plots two models for light response curves, and reports the parameters of the fits.

To Do
-implement model that accounts for 'photoinhibition' by acceleration of respiration with increasing light

# Define data file name, in .xlsx format
```{r paths and variables}
O2FileName <- "2022.03.21.Sevcik.R.Test1.xlsx"
```

# Define location of data file
```{r file path}
DataInPath <- file.path("..", "Data", "RawData")
```

# Call libraries to access functions
```{r libraries}
library(readxl) #importing files from Excel
library(tidyverse) #data wrangling & plotting
library(minpack.lm) #non linear model fitting
library(broom) #formatting model outputs
```

# Import data file into R and inspect
```{r file import}
O2Data <- read_excel(file.path(DataInPath, O2FileName))

O2Data
```
# Rename variables for R compatability
```{r variable rename}
O2Data <- O2Data %>%
  rename(umolphotons_m2_s1 = 1,
         umolO2_mmolChl1_s1 = 2)
```

# Plot Data
```{r test plot}
O2Data %>%
  ggplot() +
  geom_point(aes(x = umolphotons_m2_s1, y = umolO2_mmolChl1_s1)) +
  theme_bw()

```

# Define model equations
Citations: Ma X, Liu Q, Zhang Z, Zhang Z, Zhou Z, Jiang Y, et al. (2021) Effects of photosynthetic models on the calculation results of photosynthetic response parameters in young Larix principis- rupprechtii Mayr. plantation. PLoS ONE 16(12): e0261683. https://doi.org/10.1371/journal.pone.0261683

https://www.frontiersin.org/articles/10.3389/fpls.2020.581851/full

I: incident light, umol photons m-2 s-1
a: alpha, initial slope of light response curve under limiting light
Pmax: maximum rate of photosynthesis; umol O2 chl-1 s-1
R: dark respiration, umol O2 chl-1 s-1
b: beta 'photoinhibition' term for decline in oxygen evolution under increasing incident light
```{r define fit equations}
# simple saturating function; rectangular hyperbola
lrc <- function(I, a, Pmax, R){((a * I * Pmax)/((a*I) + Pmax)) - R
}

# function with beta photoinhibition
lrc_beta <- function(I, a, b, g, R){(a * (((1 - (b*I))/(1 + (g*I)))*I)) - R
}

# function with light-dependent Respiration
lrc_dr <- function(I, a, Pmax, R, d){((a * I * Pmax)/((a*I) + Pmax)) - (R * I * d)
}

# Solve model for light compensation point; light level when O2 from photosynthesis = O2 taken up by respiration; this is a 'breakeven point' for photosynthesis.

# 0 = (a * (((1 - (b*I))/(1 + (g*I)))*I)) - R

# R = (a * (((1 - (b*I))/(1 + (g*I)))*I))
# R/a = (((1 - (b*I))/(1 + (g*I))) * I)
# R/(a * I) = (1 - (b*I))/(1 + (g*I))

# R/a = (I - (b*I^2))/(1 + (g*I))

# R/a = (I - (b*I^2))/(1 + (g*I))

# R/a = ((I/(1 + (g*I))) - ((b*I^2)/(1 + (g*I)))

# R/a = ((I/(1 + (g*I))) - ((b*I^2)/(1 + (g*I)))

```

# Fit model equations
```{r fit models}
O2Models <- O2Data %>%
  nest(data = everything()) %>%
  mutate(LRC_model = map(data, possibly(~nlsLM(umolO2_mmolChl1_s1 ~ lrc(I = umolphotons_m2_s1, a, Pmax, R),
                              data = .x,
                              start = list(a = 1,
                                           Pmax = max(.$umolO2_mmolChl1_s1, na.rm = TRUE),
                                           R = min(.$umolO2_mmolChl1_s1, na.rm = TRUE))
                         ), NULL)
                         )
         ) %>%
  mutate(LRC_Param = map(LRC_model, tidy),
         LRC_Predict = map(LRC_model, augment)) %>%
  mutate(LRCbeta_model = map(data, possibly(~nlsLM(umolO2_mmolChl1_s1 ~ lrc_beta(I = umolphotons_m2_s1, a, b, g, R),
                              data = .x,
                              start = list(a = 1,
                                           b = 1,
                                           g = 1,
                                           R = min(.$umolO2_mmolChl1_s1, na.rm = TRUE))
                         ), NULL)
                         )
         ) %>%
  mutate(LRCbeta_Param = map(LRCbeta_model, tidy),
         LRCbeta_Predict = map(LRCbeta_model, augment)) %>%
   mutate(LRCdR_model = map(data, possibly(~nlsLM(umolO2_mmolChl1_s1 ~ lrc_dr(I = umolphotons_m2_s1, a, Pmax, R, d),
                              data = .x,
                              start = list(a = 5,
                                           Pmax = 200,
                                           R = 100,
                                           d = 0.01)
                         ), NULL)
                         )
         ) %>%
  mutate(LRCdR_Param = map(LRCdR_model, tidy),
         LRCdR_Predict = map(LRCdR_model, augment))

```

# Plot models
```{r plot models}
# Adding legend for lines
#https://community.rstudio.com/t/adding-manual-legend-to-ggplot2/41651/2

line_colours = c("umolO2_mmolChl1_s1" = "green",
            "fitted" = "black",
            "residual" = "red")

O2Models %>%
  unnest(LRC_Predict) %>%
  ggplot(aes(x = umolphotons_m2_s1)) +
  geom_point(aes(y = umolO2_mmolChl1_s1, colour =  "umolO2_mmolChl1_s1")) +
  geom_line(aes(y = `.fitted`, colour = "fitted")) +
  geom_point(aes(y = `.resid`, colour = "residual"), size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") + 
  labs(colour = "Legend") + 
  scale_colour_manual(values = line_colours) +
  scale_x_continuous(expand = c(0,0)) +
  theme_bw()

O2Models %>%
  unnest(LRCbeta_Predict) %>%
  ggplot(aes(x = umolphotons_m2_s1)) +
  geom_point(aes(y = umolO2_mmolChl1_s1, colour =  "umolO2_mmolChl1_s1")) +
  geom_line(aes(y = `.fitted`, colour = "fitted")) +
  geom_point(aes(y = `.resid`, colour = "residual"), size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") + 
  labs(colour = "Legend") + 
  scale_colour_manual(values = line_colours) +
  scale_x_continuous(expand = c(0,0)) +
  theme_bw()

# O2Models %>%
#   unnest(LRCrD_Predict) %>%
#   ggplot() +
#   geom_point(aes(x = umolphotons_m2_s1, y = umolO2_mmolChl1_s1), colour = "green") +
#   geom_line(aes(x = umolphotons_m2_s1, y = `.fitted`), colour = "black") +
#   geom_point(aes(x = umolphotons_m2_s1, y = `.resid`), colour = "red", size = 0.5) +
#   geom_hline(yintercept = 0, linetype = "dashed") + 
#   theme_bw()

```

# Report model parameters
## LRC model parameters
```{r LRC model parameters}
LRC_Param <- O2Models %>%
  unnest(LRC_Param) %>%
  select(c(term, estimate, std.error)) %>%
  mutate(Model = "lrc", before = "Value")

LRC_O2compEst <- lrc(I = seq(0,100, by = 1), a = LRC_Param$estimate[LRC_Param$term == "a"], Pmax = LRC_Param$estimate[LRC_Param$term == "Pmax"], R = LRC_Param$estimate[LRC_Param$term == "R"])

O2compEst <- which(abs(LRC_O2compEst - 0) == min(abs(LRC_O2compEst - 0)))

LRC_Param <- LRC_Param %>%
  mutate(O2comp = O2compEst)

LRC_Param2 <- tibble(Model = "lrc",
                     Value = colnames(LRC_Param)[2:3],
                     a = c(as.numeric(LRC_Param[1,2]), as.numeric(LRC_Param[1,3])),
                     Pmax = c(as.numeric(LRC_Param[2,2]), as.numeric(LRC_Param[2,3])),
                     R = c(as.numeric(LRC_Param[3,2]), as.numeric(LRC_Param[3,3]))) %>%
  mutate(Pmax_R = Pmax - R,
         O2Comp = c(O2compEst, NA))

LRC_Param2

```


## LRCbeta_model_parameters
```{r LRCbeta model parameters}
LRCbeta_Param <- O2Models %>%
  unnest(LRCbeta_Param) %>%
  select(c(term, estimate, std.error)) %>%
  mutate(Model = "lrc", before = "Value")

LRCbeta_O2compEst <- lrc_beta(I = seq(0,100, by = 1), a = LRCbeta_Param$estimate[LRCbeta_Param$term == "a"], b = LRCbeta_Param$estimate[LRCbeta_Param$term == "b"], g = LRCbeta_Param$estimate[LRCbeta_Param$term == "g"], R = LRCbeta_Param$estimate[LRCbeta_Param$term == "R"])

O2compEst_beta <- which(abs(LRCbeta_O2compEst - 0) == min(abs(LRCbeta_O2compEst - 0)))

LRCbeta_Param <- LRCbeta_Param %>%
  mutate(O2comp = O2compEst_beta)

LRCbeta_Param2 <- tibble(Model = "lrc_beta",
                         Value = colnames(LRCbeta_Param)[2:3],
                     a = c(as.numeric(LRCbeta_Param[1,2]), as.numeric(LRCbeta_Param[1,3])),
                     b = c(as.numeric(LRCbeta_Param[2,2]), as.numeric(LRCbeta_Param[2,3])),
                      g = c(as.numeric(LRCbeta_Param[3,2]), as.numeric(LRCbeta_Param[3,3])),
                     R = c(as.numeric(LRCbeta_Param[4,2]), as.numeric(LRCbeta_Param[4,3]))) %>%
  mutate(Pmax = a*((sqrt(b + g) - sqrt(b))/g)^2,
         Pmax_R = Pmax - R,
         EK = Pmax/a,
        O2Comp = c(O2compEst_beta, NA))

LRCbeta_Param2
```

# Plot including fit parameters
```{r plot parameters}
param_colours = line_colours = c("umolO2_mmolChl1_s1" = "green",
                                 "alpha" = "green",
                                 "fitted" = "black",
                                 "resid" = "red",
                                 "beta" = "orange",
                                 "Ek" = "blue",
                                 "Respiration" = "darkred",
                                 "Pmax" = "darkgreen",
                                 "O2Comp" = "red")
  
a_fun <- function(x, a = as.numeric(LRCbeta_Param2$a[1]), R = as.numeric(LRCbeta_Param2$R[1])){(x * a) - R
}

b_fun <- function(x, b = as.numeric(LRCbeta_Param2$b[1]), g = as.numeric(LRCbeta_Param2$g[1]), R = as.numeric(LRCbeta_Param2$R[1])){(x * (1-(b*x))/(1 + (g*x))) - R
}

#r_fun <- function()


O2Models %>%
  unnest(LRCbeta_Predict) %>%
  mutate(R = LRCbeta_Param2$R[1],
         PMax_R = LRCbeta_Param2$Pmax_R[1],
         Ek = LRCbeta_Param2$EK[1],
         O2Comp = LRCbeta_Param2$O2Comp[1],
         alpha = a_fun(x = umolphotons_m2_s1, a = as.numeric(LRCbeta_Param2$a[1])),
         beta = b_fun(x = umolphotons_m2_s1, b = as.numeric(LRCbeta_Param2$b[1]), g = as.numeric(LRCbeta_Param2$g[1]), R = R)) %>%
  ggplot(aes(x = umolphotons_m2_s1)) +
  xlim(0, max(as.numeric(O2Data$umolphotons_m2_s1) + 20, na.rm = TRUE)) +
  ylim(min(as.numeric(O2Data$umolO2_mmolChl1_s1), na.rm = TRUE) - 20, max(as.numeric(O2Data$umolO2_mmolChl1_s1) + 20, na.rm = TRUE)) +
  geom_point(aes(y = umolO2_mmolChl1_s1, colour =  "umolO2_mmolChl1_s1")) +
  geom_line(aes(y = `.fitted`, colour = "fitted")) +
  geom_hline(yintercept = 0, linewidth = 2) + 
  geom_line(aes(y = -R, colour = "Respiration")) +
  geom_line(aes(y = PMax_R, colour = "Pmax")) +
  geom_line(aes(x = Ek, y = umolO2_mmolChl1_s1, colour = "Ek")) +
  geom_line(aes(x = O2Comp, y = umolO2_mmolChl1_s1, colour = "O2Comp")) +
  geom_line(aes(y = alpha, colour = "alpha")) +
  geom_line(aes(y = beta, colour = "beta"), linetype = "dotted") +
  labs(colour = "Legend") + 
  scale_colour_manual(values = param_colours) +
  scale_x_continuous(expand = c(0,0)) +
  theme_bw()


# 
# 
#   geom_function(fun = a_fun, linetype = "dashed", colour = "alpha") +


 

```

