---
title: "O2 LRC Fitting"
output: html_notebook
author: Douglas A. Campbell
---

This R Notebook imports oxygen evolution vs. PAR data.
It fits and plots two models for light response curves, and reports the parameters of the fits.

# Define data file name, in .xlsx format
```{r paths and variables}
O2FileName <- "2022.03.21.Sevcik.R.Test1.xlsx"
```

# Define location of data file
```{r file path}
DataInPath <- file.path("..", "Data", "RawData")
```

# Call libraries to access functions
```{r libraries}
library(tidyverse) #data wrangling & plotting
library(minpack.lm) #non linear model fitting
library(broom) #formatting model outputs
```

# Import data file into R and inspect
```{r file import}
O2Data <- read_excel(file.path(DataInPath, O2FileName))

O2Data
```
# Rename variables for R compatability
```{r variable rename}
O2Data <- O2Data %>%
  rename(umolphotons_m2_s1 = 1,
         umolO2_mmolChl1_s1 = 2)
```

# Plot Data
```{r test plot}
O2Data %>%
  ggplot() +
  geom_point(aes(x = umolphotons_m2_s1, y = umolO2_mmolChl1_s1)) +
  theme_bw()

```

# Define model equations
Citation: Ma X, Liu Q, Zhang Z, Zhang Z, Zhou Z, Jiang Y, et al. (2021) Effects of photosynthetic
models on the calculation results of photosynthetic response parameters in young Larix principis-
rupprechtii Mayr. plantation. PLoS ONE 16(12): e0261683. https://doi.org/10.1371/journal.pone.0261683

I: incident light, umol photons m-2 s-1
a: alpha, initial slope of light response curve under limiting light
Pmax: maximum rate of photosynthesis; umol O2 chl-1 s-1
R: dark respiration, umol O2 chl-1 s-1
b: beta 'photoinhibition' term for decline in oxygen evolution under increasing incident light
```{r define fit equations}
# simple saturating function; rectangular hyperbola
lrc <- function(I, a, Pmax, R){((a * I * Pmax)/((a*I) + Pmax)) - R
}

# function with beta photoinhibition
lrc_beta <- function(I, a, b, g, R){(a * (((1 - (b*I))/(1 + (g*I)))*I)) - R
}

# function with beta photoinhibition
lrc_beta_alt <- function(I, a, b, Pmax, R){(a * (((1 - (b*I))/(1 + ((a/Pmax)*I)))*I)) - R
}


```

# Fit model equations
```{r fit models}
O2Models <- O2Data %>%
  nest(data = everything()) %>%
  mutate(LRC_model = map(data, possibly(~nlsLM(umolO2_mmolChl1_s1 ~ lrc(I = umolphotons_m2_s1, a, Pmax, R),
                              data = .x,
                              start = list(a = 1,
                                           Pmax = max(.$umolO2_mmolChl1_s1, na.rm = TRUE),
                                           R = min(.$umolO2_mmolChl1_s1, na.rm = TRUE))
                         ), NULL)
                         )
         ) %>%
  mutate(LRC_Param = map(LRC_model, tidy),
         LRC_Predict = map(LRC_model, augment)) %>%
  mutate(LRCbeta_model = map(data, possibly(~nlsLM(umolO2_mmolChl1_s1 ~ lrc_beta(I = umolphotons_m2_s1, a, b, g, R),
                              data = .x,
                              start = list(a = 1,
                                           b = 1,
                                           g = 1,
                                           R = min(.$umolO2_mmolChl1_s1, na.rm = TRUE))
                         ), NULL)
                         )
         ) %>%
  mutate(LRCbeta_Param = map(LRCbeta_model, tidy),
         LRCbeta_Predict = map(LRCbeta_model, augment))

```

# Plot models
```{r plot models}
O2Models %>%
  unnest(LRC_Predict) %>%
  ggplot() +
  geom_point(aes(x = umolphotons_m2_s1, y = umolO2_mmolChl1_s1), colour = "green") +
  geom_line(aes(x = umolphotons_m2_s1, y = `.fitted`), colour = "black") +
  geom_point(aes(x = umolphotons_m2_s1, y = `.resid`), colour = "red", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") + 
  theme_bw()

O2Models %>%
  unnest(LRCbeta_Predict) %>%
  ggplot() +
  geom_point(aes(x = umolphotons_m2_s1, y = umolO2_mmolChl1_s1), colour = "green") +
  geom_line(aes(x = umolphotons_m2_s1, y = `.fitted`), colour = "black") +
  geom_point(aes(x = umolphotons_m2_s1, y = `.resid`), colour = "red", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") + 
  theme_bw()

```

# Report model parameters
```{r model parameters}
LRC_Param <- O2Models %>%
  unnest(LRC_Param) %>%
  select(c(term, estimate, std.error))

LRC_Param


# LRC_Param %>%
#   pivot_longer()
#   pivot_wider(names_from = term, values_from = c(estimate, std.error))

LRCbeta_Param <- O2Models %>%
  unnest(LRCbeta_Param) %>%
  select(c(term, estimate, std.error))
  

LRCbeta_Param %>%
  pivot_wider(names_from = term, values_from = c(estimate))
```


